---
title: "Comparisons of Differential Expression Tools on Different Data Sources"
output: html_notebook
---

# Setup
```{r}
library(biomaRt)
library(DESeq2)
library(sleuth)
library(tximport)
library(jsonlite)
library(data.table)
library(vsn)
library(EnhancedVolcano)
library(viridis)
library(ggpubr)
library(MASS)
library(rtracklayer)
library(VennDiagram)
library(RColorBrewer)
library(ComplexHeatmap)
library(RColorBrewer)
library(circlize)
library(tidyverse)
set.seed(42)
RESPONSE_COLORS = c("NR"="purple","R"="orange")
```

## Setup sample details
```{r}
# Load in samples and point to kallisto paths
samples <- read_tsv("../work/input/sample_manifest.tsv")
samples$kallisto_path <- file.path("..","work","intermediate","quant",samples$SRR)
```

## Transcript <-> Gene Mappings
```{r}
mart <- biomaRt::useMart(biomart = "ENSEMBL_MART_ENSEMBL",
  dataset = "hsapiens_gene_ensembl",
  host = 'grch37.ensembl.org')
gene_features <- biomaRt::getBM(attributes = c("ensembl_transcript_id_version", "ensembl_gene_id",
    "external_gene_name","transcript_length"), mart = mart)
t2g <- gene_features %>%
  select(ensembl_transcript_id_version,ensembl_gene_id,external_gene_name)
avg_tx_len <-gene_features %>%
  group_by(external_gene_name) %>%
  summarise(length=mean(transcript_length))
```

## QC Filtering
We'll filter high quality samples by percent reads mapped by Kallisto. We'll need to get hte total reads in files from the kallisto logs, and the total number of mapped reads by pre-loading all the counts using sleuth. 
```{r}
# Find total reads to compute percent mapped
samples$n_reads <- apply(samples,1,function(x)read_json(file.path(x["kallisto_path"],"run_info.json"))$n_processed)
```

```{r}
# Initial sleuth just to get kallisto counts
s_obj_counts <- sleuth_prep(
  sample_to_covariates = rename(samples,path=kallisto_path,sample=SRR),
  normalize=FALSE
)
```

```{r}
kallisto_est_mapped <- s_obj_counts$obs_raw %>%
  group_by(sample) %>%
  summarise(kallisto_mapped_reads=sum(est_counts))
samples <- samples %>% 
  left_join(kallisto_est_mapped,by=c("SRR"="sample")) %>%
  mutate(percent_mapped=kallisto_mapped_reads/n_reads) 
samples %>%
  ggplot(aes(x = percent_mapped)) +
    geom_histogram(bins = 10) +
    labs(title = "Distribution of % Mapped Reads", x = "% Mapped Reads")
```

```{r}
# Filter samples
samples_filt <-samples %>% 
  filter(
    !SRR %in% c("SRR8526726"),
    #percent_mapped > 0.8
    )
```



# Sleuth
```{r include=FALSE}
# Setup sleuth object
s_obj <- sleuth_prep(
  sample_to_covariates = rename(samples_filt,path=kallisto_path,sample=SRR),
  full_model = ~Response,
  transform_fun_counts=function(x)log2(x+1),
  target_mapping = rename(t2g,target_id=ensembl_transcript_id_version) %>%
      distinct(target_id,external_gene_name),
  aggregation_column = "external_gene_name",
  gene_mode=TRUE
)
```

## PCA 

### Sleuth default
```{r}
plot_pca(s_obj,units="scaled_reads_per_base",color_by = "Response",use_filtered = TRUE,text_labels = F) + geom_label(aes(label=sample),size=2)
```

### Standardized PCA
```{r}

sleuth_rna_norm <- s_obj$obs_norm_filt %>% 
  select(sample,target_id,scaled_reads_per_base) %>%
  spread(key=target_id,value=scaled_reads_per_base) %>%
  column_to_rownames("sample")
sleuth_std_pca <- prcomp((sleuth_rna_norm),center=T,scale.=T)
sleuth_variance_explained = summary(sleuth_std_pca)$importance[2,1:2]
sleuth_std_pca$x[,1:2] %>%
  as.data.frame() %>%
  rownames_to_column("SRR") %>%
  left_join(samples,by="SRR") %>%
  ggplot(aes(x=PC1,y=PC2,color=Response,label=SRR)) +
    geom_point(size=3) +
    labs(title="PCA Plot of Sleuth Normalized Kallisto Quantified Data",
         x = paste0("PC1: ", round(sleuth_variance_explained[1]*100),"% Variance"),
         y = paste0("PC2: ", round(sleuth_variance_explained[2]*100),"% Variance")) +
    scale_color_manual(values=RESPONSE_COLORS)
```

## Diffex

### LRT
```{r}
s_obj <- s_obj %>%
  sleuth_fit(~Response,"full") %>%
  sleuth_fit(~1,"reduced") %>%
  sleuth_lrt("reduced","full")
```
```{r}
sleuth_diffex_lrt <- sleuth_results(s_obj,test="reduced:full",test_type="lrt") %>% filter(!is.na(pval))
sleuth_diffex_lrt %>% arrange(qval)
```

### Wald
```{r}
s_obj <- sleuth_wt(s_obj,"ResponseR")
sleuth_diffex_wald <- sleuth_results(s_obj,test="ResponseR",test_type="wald")  %>% filter(!is.na(pval))
sleuth_diffex_wald %>% arrange(qval) 
```
```{r}
plot_volcano(s_obj,test_type="wt",test="ResponseR",sig_level=0.05) +
  labs(title="Volcano Plot of Sleuth Identified Differentially Expressed Genes",
       x="log2FC", y = "qval",
        color="Q < 0.05")
```

## QC
```{r}
meanSdPlot(log2(sleuth_to_matrix(s_obj,"obs_norm","scaled_reads_per_base") + 1),rank=FALSE,bins=100)
```

# DESeq (Kallisto)
```{r}
files <- file.path(samples_filt$kallisto_path,"abundance.h5")
names(files) <- samples_filt$`Patient ID`
txi <- tximport(
  file.path(samples_filt$kallisto_path,"abundance.h5"),
  type = "kallisto",
  tx2gene = select(
    t2g,
    TXNAME = ensembl_transcript_id_version,
    GENEID = external_gene_name
  ),
  ignoreAfterBar = TRUE,
)
rownames(samples_filt) <- samples_filt$`Patient ID`
deseq_kallisto_data <- DESeqDataSetFromTximport(txi,samples_filt,~Response)
kallisto_keep <- rowSums(counts(deseq_kallisto_data)>5)>5
deseq_kallisto_data <- deseq_kallisto_data[kallisto_keep,]
```

## QC
```{r}
deseq_kallisto_norm <- vst(deseq_kallisto_data)
meanSdPlot(assay(deseq_kallisto_norm),ranks=FALSE,bins=100)

```


## PCA
```{r}
plotPCA(deseq_kallisto_norm,intgroup="Response",ntop=500000)   +
  labs(title="PCA Plot of DESeq Normalized Kallisto Quantified Data",
       color="Response") +
  scale_color_manual(values=RESPONSE_COLORS)
```
## Difex
```{r}
deseq_kallisto_data <- DESeq(deseq_kallisto_data)
deseq_kallisto_lrt <- results(deseq_kallisto_data,name="Response_R_vs_NR",tidy=TRUE)
deseq_kallisto_lrt %>% 
  #filter(padj < 0.05) %>%
  arrange(padj) %>%
  select(row,log2FoldChange,pvalue,padj) %>%
  left_join(select(t2g,ensembl_gene_id,external_gene_name),by=c("row"="ensembl_gene_id"))
```

# DESeq (FeatureCounts)
```{r}
fc_mat <- read_tsv("../work/input/GSE126044_counts.txt")
 fc_mat <- fc_mat[!grepl("^\\d+\\-",fc_mat$X1),] %>%
   column_to_rownames("X1")
 
fc_mat_filt <-  fc_mat %>%
   select(one_of(samples_filt$`Patient ID`))
deseq_fc <- DESeqDataSetFromMatrix(countData=fc_mat_filt,colData=samples_filt,design=~Response)
fc_keep <- rowSums(counts(deseq_fc)>5)>5
deseq_fc <- deseq_fc[fc_keep,]
```

## QC
```{r}
deseq_fc_norm <- vst(deseq_fc)
meanSdPlot(assay(deseq_fc_norm),rank=FALSE,bins=100)
```


## PCA
```{r}
plotPCA(deseq_fc_norm,intgroup="Response",ntop=500000) +
  labs(title="PCA Plot of DESeq Normalized Feature Counts/STAR Quantified Data",
       color="Response") +
  scale_color_manual(values=RESPONSE_COLORS)
```

## Difex
```{r}
deseq_fc <- DESeq(deseq_fc)
deseq_fc_lrt <- results(deseq_fc,name="Response_R_vs_NR",tidy=TRUE)
deseq_fc_lrt %>% 
 # filter(padj < 0.05) %>%
  arrange(padj) %>%
  select(row,log2FoldChange,pvalue,padj)
```

# Merged Analyses

## counts_to_rpm, pca
```{r}
tfc <- t(fc_mat_filt)
# Convert Kallisto counts to rpm
kalread<-cbind(kallisto_mapped_reads=samples_filt$kallisto_mapped_reads, sleuth_rna_norm %>% rownames_to_column("sample"))
kalrpm<-setDT(kalread) [, lapply(.SD, function(x) (x * 1e6) / (kallisto_mapped_reads)),  by=.(kallisto_mapped_reads,sample)] %>% 
  column_to_rownames("sample")

# Convert star fc to rpm
#get df with est total reads for star
staread<-cbind(star_est_mapped=colSums(fc_mat_filt), as.data.frame(tfc) %>% rownames_to_column("sample"))
starpm<-setDT(staread) [, lapply(.SD, function(x) (x * 1e6) / (star_est_mapped)),  by=.(star_est_mapped,sample)] %>%
  column_to_rownames("sample")

## scale pca with rpm
kal_rpm_pca <- prcomp(kalrpm[,-1],center=TRUE,scale.=TRUE)

starnorm_std <- scale(starpm[,-1])
#star counts include some zero, set scale/std apply to those !=0
star_rpm_pca <- prcomp(starnorm_std[ , which(apply(starnorm_std, 2, var) != 0)],center=TRUE)

## superimposed pca plots from the two sources
p1 <-kal_rpm_pca$x[,1:2]%>%
  as.data.frame() %>%
  rownames_to_column("SRR") %>%
  left_join(samples_filt,by="SRR")
p2<-star_rpm_pca$x[,1:2] %>%
  as.data.frame() %>%
  rownames_to_column("Patient ID") %>%
  left_join(samples_filt,by="Patient ID")
bind_rows(Kallisto=p1,STAR=p2,.id="quant_type") %>%
  ggplot(aes(x=PC1,y=PC2,color=Response,shape=quant_type)) +
    geom_point(size=3) +
    scale_color_manual(values=RESPONSE_COLORS) + 
    labs(title="Superimposed PCA of RPM Normalized Counts",
         shape="Method")
```



## Comparison of log2FC
```{r}
equation = function(x) {
  lm_coef <- list(a = round(as.numeric(coef(x)[1]), digits = 2),
                  b = round(as.numeric(coef(x)[2]), digits = 2),
                  r2 = round(summary(x)$r.squared, digits = 2));
  lm_eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(R)^2~"="~r2,lm_coef)
  as.character(as.expression(lm_eq));                 
}

log2fc_combined <- sleuth_diffex_wald %>%
  filter(!is.na(b)) %>%
  select(gene=target_id,sleuth_log2FC=b,sleuth_q=qval) %>%
  left_join(select(deseq_kallisto_lrt,row,deseq_kal_log2FC=log2FoldChange,deseq_kal_q=padj),by=c("gene"="row")) %>%
  left_join(select(deseq_fc_lrt,row,deseq_fc_log2FC=log2FoldChange,deseq_fc_q=padj),by=c("gene"="row")) %>%
  filter(!is.na(deseq_fc_log2FC))

regression.line <- lm(deseq_kal_log2FC ~ sleuth_log2FC, data = log2fc_combined)
log2fc_combined %>%
  ggplot(aes(x=sleuth_log2FC,y=deseq_kal_log2FC)) +
  geom_point(aes(x=sleuth_log2FC,y=deseq_kal_log2FC)) +
  geom_hline(yintercept=0, color="darkgray") + 
  geom_vline(xintercept=0, color="darkgray") +
  theme_pubr(base_size = 14)+
  geom_smooth(method = "lm", color="red") +
  annotate("text", label = equation(regression.line), x = -3, y = 4, parse = TRUE)+
  labs(title="Comparison of log2FC between Sleuth and DESeq Kallisto",
       x="log2 Fold Change (Sleuth-Kallisto)",y="log2 Fold Change (DESeq-Kallisto)")


regression.line <- lm(deseq_fc_log2FC ~ deseq_kal_log2FC, data = log2fc_combined)
log2fc_combined %>%
  ggplot(aes(x=deseq_kal_log2FC,y=deseq_fc_log2FC)) +
  geom_point() +
  geom_hline(yintercept=0, color="darkgray") + 
  geom_vline(xintercept=0, color="darkgray") +
  theme_pubr(base_size = 14)+
  geom_smooth(method = "lm", color="red") +
  annotate("text", label = equation(regression.line), x = -4, y = 5, parse = TRUE)+
  labs(title="Comparison of log2FC between DESeq From Kallisto and FC",
       x="log2 Fold Change (DESeq-FeatureCounts)",y="log2 Fold Change (DESeq-Kallisto)")

```



## combine all logFC data
```{r}
sleuth_diffex_wald <- sleuth_diffex_wald %>% filter(!is.na(b)) 
deseq_fc_lrt <- deseq_fc_lrt %>% filter(!is.na(log2FoldChange)) 
deseq_kallisto_lrt <- deseq_kallisto_lrt %>% filter(!is.na(log2FoldChange)) 

log2fc_full_combined <- sleuth_diffex_wald %>%
  select(gene=target_id,sleuth_log2FC=b,sleuth_q=qval) %>%
  full_join(select(deseq_kallisto_lrt,row,deseq_kal_log2FC=log2FoldChange,deseq_kal_q=padj),by=c("gene"="row")) %>%
  full_join(select(deseq_fc_lrt,row,deseq_fc_log2FC=log2FoldChange,deseq_fc_q=padj),by=c("gene"="row"))
log2fc_full_combined

```

## Volcano Plots

### Slueth
```{r}

log2fc_full_combined$log2FoldChange <- log2fc_full_combined$sleuth_log2FC
log2fc_full_combined$padj <- log2fc_full_combined$sleuth_q
log2fc_final <- log2fc_full_combined %>% filter(!is.na(log2FoldChange)) 

thres <- 1
keyvals.color <- rep('green', nrow(log2fc_final))
# set the base name/label as 'Mid'
names(keyvals.color) <- rep('Mid', nrow(log2fc_final))

keyvals.color[which(log2fc_final$log2FoldChange > thres)] <- 'red'
names(keyvals.color)[which(log2fc_final$log2FoldChange > thres)] <- 'UpRegulated'

keyvals.color[which(log2fc_final$log2FoldChange < -thres)] <- 'royalblue'
names(keyvals.color)[which(log2fc_final$log2FoldChange < -thres)] <- 'DownRegulated'

down.genes  <- which((log2fc_final$padj < 0.3) & (log2fc_final$log2FoldChange <= -3))
up.genes  <- which((log2fc_final$padj < 0.3) & (log2fc_final$log2FoldChange >= 3))
up.down.genes = c(down.genes, up.genes)
length(up.down.genes)

table(keyvals.color)
EnhancedVolcano(log2fc_final,
                lab = (log2fc_final$gene),
                x = 'log2FoldChange',
                y = 'padj', 
                title = 'R vs. NR  - sleuth',
                drawConnectors = TRUE,
                widthConnectors = 0.3,
                colConnectors = 'grey30',
                colCustom = keyvals.color,
                pCutoff = NA,
                cutoffLineType = 'blank',
                vline = c(-thres,thres),
                vlineCol = c('grey50', 'grey50'),
                gridlines.major = FALSE,
                gridlines.minor = FALSE,
                ylim = c(0,1.5),
                selectLab = log2fc_final$gene[up.down.genes]
                )

```

### deseq_kal
```{r}

log2fc_full_combined$log2FoldChange <- log2fc_full_combined$deseq_kal_log2FC
log2fc_full_combined$padj <- log2fc_full_combined$deseq_kal_q
log2fc_final <- log2fc_full_combined %>% filter(!is.na(log2FoldChange)) 

thres <- 1
keyvals.color <- rep('green', nrow(log2fc_final))
# set the base name/label as 'Mid'
names(keyvals.color) <- rep('Mid', nrow(log2fc_final))

keyvals.color[which(log2fc_final$log2FoldChange > thres)] <- 'red'
names(keyvals.color)[which(log2fc_final$log2FoldChange > thres)] <- 'UpRegulated'

keyvals.color[which(log2fc_final$log2FoldChange < -thres)] <- 'royalblue'
names(keyvals.color)[which(log2fc_final$log2FoldChange < -thres)] <- 'DownRegulated'

down.genes  <- which((log2fc_final$padj < 0.3) & (log2fc_final$log2FoldChange <= -20))
up.genes  <- which((log2fc_final$padj < 0.3) & (log2fc_final$log2FoldChange >= 5))
up.down.genes = c(down.genes, up.genes)
length(up.down.genes)

table(keyvals.color)
EnhancedVolcano(log2fc_final,
                lab = (log2fc_final$gene),
                x = 'log2FoldChange',
                y = 'padj', 
                title = 'R vs. NR  - Deseq-kallisto',
                drawConnectors = TRUE,
                widthConnectors = 0.3,
                colConnectors = 'grey30',
                colCustom = keyvals.color,
                pCutoff = NA,
                cutoffLineType = 'blank',
                vline = c(-thres,thres),
                vlineCol = c('grey50', 'grey50'),
                gridlines.major = FALSE,
                gridlines.minor = FALSE,
                ylim = c(0,10),
                selectLab = log2fc_final$gene[up.down.genes]
                )

```


### deseq_fc
```{r}
log2fc_full_combined$log2FoldChange <- log2fc_full_combined$deseq_fc_log2FC
log2fc_full_combined$padj <- log2fc_full_combined$deseq_fc_q
log2fc_final <- log2fc_full_combined %>% filter(!is.na(log2FoldChange)) 

thres <- 1
keyvals.color <- rep('green', nrow(log2fc_final))
# set the base name/label as 'Mid'
names(keyvals.color) <- rep('Mid', nrow(log2fc_final))

keyvals.color[which(log2fc_final$log2FoldChange > thres)] <- 'red'
names(keyvals.color)[which(log2fc_final$log2FoldChange > thres)] <- 'UpRegulated'

keyvals.color[which(log2fc_final$log2FoldChange < -thres)] <- 'royalblue'
names(keyvals.color)[which(log2fc_final$log2FoldChange < -thres)] <- 'DownRegulated'

down.genes  <- which((log2fc_final$padj < 0.3) & (log2fc_final$log2FoldChange <= -8))
up.genes  <- which((log2fc_final$padj < 0.3) & (log2fc_final$log2FoldChange >= 4))
up.down.genes = c(down.genes, up.genes)
length(up.down.genes)

table(keyvals.color)
EnhancedVolcano(log2fc_final,
                lab = (log2fc_final$gene),
                x = 'log2FoldChange',
                y = 'padj', 
                title = 'R vs. NR  - Deseq-fc',
                drawConnectors = TRUE,
                widthConnectors = 0.3,
                colConnectors = 'grey30',
                colCustom = keyvals.color,
                pCutoff = NA,
                cutoffLineType = 'blank',
                vline = c(-thres,thres),
                vlineCol = c('grey50', 'grey50'),
                gridlines.major = FALSE,
                gridlines.minor = FALSE,
                ylim = c(0,15),
                selectLab = log2fc_final$gene[up.down.genes]
                )

```


## Venn diagram
```{r}
thres <- 1
# get the name of the upregulated/downregulated genes from 3 different methods
sleuth.up <- sleuth_diffex_wald$target_id[which((sleuth_diffex_wald$qval <= 0.05) & (sleuth_diffex_wald$b >= thres))]
sleuth.down <- sleuth_diffex_wald$target_id[which((sleuth_diffex_wald$qval <= 0.05) & (sleuth_diffex_wald$b <= -thres))]
deseq.k.up <- deseq_kallisto_lrt$row[which((deseq_kallisto_lrt$padj <= 0.05) & (deseq_kallisto_lrt$log2FoldChange >= thres))]
deseq.k.down <- deseq_kallisto_lrt$row[which((deseq_kallisto_lrt$padj <= 0.05) & (deseq_kallisto_lrt$log2FoldChange <= -thres))]
deseq.fc.up <- deseq_fc_lrt$row[which((deseq_fc_lrt$padj <= 0.05) & (deseq_fc_lrt$log2FoldChange >= thres))]
deseq.fc.down <- deseq_fc_lrt$row[which((deseq_fc_lrt$padj <= 0.05) & (deseq_fc_lrt$log2FoldChange <= -thres))]

# up-regulated venn-diagram
grid.newpage()
vd <- venn.diagram(x = list("Sleuth-Kallisto" = (sleuth.up),
                            "DESeq-Kallisto" = (deseq.k.up),
                            "DESeq-FeatureCounts" = (deseq.fc.up)),
                    main="Venn Diagram of Upregulated Genes Between 3 Methods",
                   fill = brewer.pal(4, "Set2")[1:3], filename = NULL)
grid.draw(vd)

# down-regulated venn-diagram
grid.newpage()
vd <- venn.diagram(x = list("Sleuth-Kallisto" = (sleuth.down),
                            "DESeq-Kallisto" = (deseq.k.down),
                            "DESeq-FeatureCounts" = (deseq.fc.down)),
                   main="Venn Diagram of Downregulated Genes Between 3 Methods",
                   fill = brewer.pal(4, "Set2")[1:3], filename = NULL)
grid.draw(vd)

```


## Heatmap
```{r}

#sleuth.up <- sleuth_diffex_wald$target_id[which((sleuth_diffex_wald$qval <= 0.05) & (sleuth_diffex_wald$b >= thres))]
#sleuth.down <- sleuth_diffex_wald$target_id[which((sleuth_diffex_wald$qval <= 0.05) & (sleuth_diffex_wald$b <= -thres))]

normalized_counts <- counts(deseq_kallisto_data, normalized=TRUE)
colnames(normalized_counts)


updown.gene <- deseq_kallisto_lrt[which((deseq_kallisto_lrt$log2FoldChange > 1 | deseq_kallisto_lrt$log2FoldChange < -1) & deseq_kallisto_lrt$padj < 0.05),]
updown.gene$updown <- ifelse((updown.gene$log2FoldChange> 1), "up", "down")

updown.regul <- as.matrix(updown.gene[,"updown", drop=FALSE])

heatmapdata <- normalized_counts[match(updown.gene$row, row.names(normalized_counts)),]

zscore_data <- t(apply(heatmapdata, 1, scale))
colnames(zscore_data) = colnames(heatmapdata)

col_updown = c("up" = "pink", "down" = "purple")

row.names(samples_filt) <- samples_filt$`Patient ID`
samples_filt

ha = HeatmapAnnotation(type = samples_filt$Response, annotation_name_side = "left",  
                       col = list(type = c("NR" = "yellow", "R" = "black")))
dim(normalized_counts)

ht1 = Heatmap(zscore_data, name = "expression",
        col = colorRamp2(c(-2, 0, 2), c("blue", "white", "red")),
        show_column_names = FALSE, show_row_names=FALSE,
        row_title = NULL,
        bottom_annotation = ha,
        heatmap_width = 400, heatmap_height = 100)

ht2 = Heatmap(updown.regul, name = "up vs down", col = col_updown, show_row_names=FALSE)
ht_list = ht1 + ht2
draw(ht_list, row_km = 1, row_split = updown.regul, cluster_rows = TRUE)
```

```{r}

#sleuth.up <- sleuth_diffex_wald$target_id[which((sleuth_diffex_wald$qval <= 0.05) & (sleuth_diffex_wald$b >= thres))]
#sleuth.down <- sleuth_diffex_wald$target_id[which((sleuth_diffex_wald$qval <= 0.05) & (sleuth_diffex_wald$b <= -thres))]
normalized_counts <- counts(deseq_fc, normalized=TRUE)

updown.gene <- deseq_fc_lrt[which((deseq_fc_lrt$log2FoldChange > 1 | deseq_fc_lrt$log2FoldChange < -1) & deseq_fc_lrt$padj < 0.05),]
updown.gene$updown <- ifelse((updown.gene$log2FoldChange> 1), "up", "down")

updown.regul <- as.matrix(updown.gene[,"updown", drop=FALSE])

heatmapdata <- normalized_counts[match(updown.gene$row, row.names(normalized_counts)),]

zscore_data <- t(apply(heatmapdata, 1, scale))
colnames(zscore_data) = colnames(heatmapdata)

col_updown = c("up" = "pink", "down" = "purple")

row.names(samples_filt) <- samples_filt$`Patient ID`
samples_filt

ha = HeatmapAnnotation(type = samples_filt$Response, annotation_name_side = "left",  
                       col = list(type = c("NR" = "yellow", "R" = "black")))
dim(normalized_counts)

ht1 = Heatmap(zscore_data, name = "expression",
        col = colorRamp2(c(-2, 0, 2), c("blue", "white", "red")),
        show_column_names = FALSE, show_row_names=FALSE,
        row_title = NULL,
        bottom_annotation = ha,
        heatmap_width = 400, heatmap_height = 100)

ht2 = Heatmap(updown.regul, name = "up vs down", col = col_updown, show_row_names=FALSE)
ht_list = ht1 + ht2
draw(ht_list, row_km = 1, row_split = updown.regul, cluster_rows = TRUE)
```

# Sample overview table
```{r}
sample_summary <- data.frame(fc_mapped_reads=colSums(fc_mat)) %>% 
  rownames_to_column("Patient ID") %>% 
  left_join(samples,by="Patient ID") %>%
  select(`Patient ID`,SRR,Response,"Sequenced Reads"=n_reads,"Mapped Reads(Kallisto)"=kallisto_mapped_reads,"Mapped Reads(STAR/FeatureCounts)"=fc_mapped_reads)
sample_summary
```


# Write out relevant files
```{r}
sleuth_diffex_lrt %>% write_tsv("sleuth_diffex_lrt.tsv")
sleuth_diffex_wald %>% write_tsv("sleuth_diffex_wald.tsv")
deseq_kallisto_lrt %>% write_tsv("deseq_kallisto_lrt.tsv")
deseq_fc_lrt %>% write_tsv("deseq_fc_lrt.tsv")
sample_summary%>% write_tsv("sample_summary.tsv")
```
